---
title: "Penguin Size Comparisons"
author: "Lindsay Hafen"
date: "11/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests
-Compare means between 2 groups - enough evidence to make us think that the data is drawn from populaitons with different means?
- Mass comparison between male and female chinstrap penguins

- Could filter male and females to make histograms, but ggplot can do that for us with the facet capability

- QQ plot compare our sample to data if it were perfectly drawn from a population with perfectly normal distribution (closer to straight line = closer to normal dist)

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

# Histograms
ggplot(chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10, fill = "darkmagenta", color = "black") +
  facet_wrap(~ sex)

## QQ plot:
ggplot(chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~ sex)

# Descriptive statistics 
chinstraps %>% 
  group_by(sex) %>% 
  summarize(mean_mass = mean(body_mass_g),
            median_mass = median(body_mass_g),
            sd_mass = sd(body_mass_g),
            sample_size = n())
  


```

